---
title: "Testing contour plots of CTD-data from ?KOKYST Klima"
author: "Helene Frigstad"
date: "3 oktober 2019"
output: html_document
---

## 0. Libraries
```{r error = F, warning = F}
#install.packages("chron")
library(tidyverse)
library(chron)
library (lubridate)
library(ggplot2)
#install.packages("akima")
library(akima)  # aspline and interp (interpolation)
library(mgcv)
#install.packages("fields")
#library(fields)  # image.plot
#library(scales) # for muted colors
```

## 1. Read data

```{r error = F, warning = F}
#clear workspace
#ls()
rm(list = ls())

# dir()
# df <- read.csv("Test_data_Helene/OKOKYST_Klima_Oct19.csv")
df <- read.csv("OKOKYST_Klima_Jun19.csv", stringsAsFactors = FALSE)
# df <- read.csv("Test_data_Helene/OKOKYST_Klima_Oct19.csv")

# convert to Date format (read as factor for some reason...) 
df$Date <- as.Date(df$Date)

```

## 2. Make datasets
```{r}

glimpse (df)

df <- df %>%
  rename(Temperature = Temperatur,
         Salinity = Salinitet,
         Oxygen = O2_ml,
         Chla = KlfA)

# collapse on Depth1 
df_m <- df %>%
group_by(Date, StationId, Depth1) %>%
  summarize_at(vars(Temperature:DIN), mean, na.rm = TRUE)

#range (df_m$O2_sat, na.rm = T)

# check O2_sat
# df_m %>%
#   #filter(StationId == "VT3") %>%
#   ggplot(aes(Date,O2_sat)) +
#     geom_line(aes()) +
#     geom_point() +
#   facet_grid(StationId~.) +
#     labs(x = "", y = "O2 sat") +
#   scale_x_date(date_breaks = "2 month", date_labels =  "%b %y") +
# theme(axis.text.x=element_text(angle=60, hjust=1))

# create dataset for VT3
VT3 <- df_m %>%
  filter(StationId == "VT3") %>%
  ungroup() %>%
  select (Date, Depth1, Temperature:DIN) %>%
  mutate(Depth = Depth1 * -1)

# create dataset for VR54
VR54 <- df_m %>%
  filter(StationId == "VR54") %>%
  ungroup() %>%
  select (Date, Depth1, Temperature:DIN) %>%
  mutate(Depth = Depth1 * -1)

# create dataset for VT71
VT71 <- df_m %>%
  filter(StationId == "VT71") %>%
  ungroup() %>%
  select (Date, Depth1, Temperature:DIN) %>%
  mutate(Depth = Depth1 * -1)

```


## 3. Preparation for plotting
* Using tutorial: https://stackoverflow.com/questions/39798500/depth-profiling-visualization

### a. Creating Matlab jet colormap
```{r}
# define a color palette similar to Matlabs jet colormap
jet.colors <- colorRampPalette(c("#00007F", "blue", "#007FFF", "cyan", "#7FFF7F", "yellow", "#FF7F00", "red", "#7F0000"))

# from Andrew:
# function J = jetak(m)
# %JET    Variant of HSV
# %   JET(M) returns an M-by-3 matrix containing the jet colormap, a variant
# %   of HSV(M). The colors begin with dark blue, range through shades of
# %   blue, cyan, green, yellow and red, and end with dark red. JET, by
# %   itself, is the same length as the current figure's colormap. If no
# %   figure exists, MATLAB uses the length of the default colormap.

```

## 4. Create contourplots

### VT3
#### a. Temperature
```{r}
# think about: 
# set the time period to equal length? or make a ggplot theme that shows avaiable data, but keeps the xlim range the same... CTD data mostly Mar 17 to nov 18
# arrange one variable in a panel for all three stations? or several vars? se ggplot tutorial for package

df_raw <- VT3  %>%
  select (Date, Depth, Temperature) %>%
  rename (Variable = Temperature) %>%
  drop_na()

# Interpolate using GAM with tensor product smooth (te)
mod1 <- gam(Variable ~ te(as.numeric(Date), Depth), data = df_raw)
mod2 <- gam(Variable ~ te(as.numeric(Date), Depth, k = 20), data = df_raw)

# Make data for predicting values (every 1 day and every 1 meter)
df_pred <- expand.grid(
  list(
    Date = seq(min(df_raw$Date), max(df_raw$Date), by = "days"),
    Depth = seq(min(df_raw$Depth), max(df_raw$Depth))))

# Use models to predict
df_pred$Var_gam1 <- predict(mod1, df_pred)
df_pred$Var_gam2 <- predict(mod2, df_pred)

# Upper 30m only
raw_30 <- df_raw %>%
  filter (Depth >= -30) 
pred_30 <- df_pred %>%
  filter (Depth >= -30) 

# whole water column
Temp_VT3 <- ggplot(df_pred, aes(x = Date, y = Depth)) + 
  geom_raster(aes(fill = Var_gam2), alpha = 0.5) +            # interpolation
    scale_fill_gradientn(colours = jet.colors(20), limits = c(-2, 25)) + 
  geom_point(data = df_raw, aes(colour = Variable), shape = "-", size = rel(6)) +  # raw
    scale_colour_gradientn(colours = jet.colors(20), limits = c(-2, 25), guide = 'none') +
  geom_contour(data = df_pred, aes(z = Var_gam2), color = "white", alpha = 0.5) +
      labs(title = "VT3", x = "", y = "Temperature (C)", fill = "") +
      scale_x_date(date_breaks = "2 month", date_labels =  "%b %y") +
      theme(axis.text.x=element_text(angle=60, hjust=1), plot.title = element_text(hjust = 0.5))

plot (Temp_VT3)
#?ggsave
#ggsave ("Figures/_Nov19/Contour_VT3_Temp.pdf", dpi=300)

# upper 30m
Temp_VT3_30 <- ggplot(pred_30, aes(x = Date, y = Depth)) + 
  geom_raster(aes(fill = Var_gam2), alpha = 0.7) +            # interpolation
    scale_fill_gradientn(colours = jet.colors(20), limits = c(-2, 25)) + 
  geom_point(data = raw_30, aes(colour = Variable), shape = "-", size = rel(6)) +  # raw
    scale_colour_gradientn(colours = jet.colors(20), limits = c(-2, 25), guide = 'none') +
  geom_contour(data = pred_30, aes(z = Var_gam2), color = "white", alpha = 0.5) +
      labs(title = "VT3", x = "", y = "Temperature (C)", fill = "") +
      scale_x_date(date_breaks = "2 month", date_labels =  "%b %y") +
      theme(axis.text.x=element_text(angle=60, hjust=1), plot.title = element_text(hjust = 0.5))

plot (Temp_VT3_30)
#ggsave ("Figures/_Nov19/Contour_VT3_Temp_30m.pdf", width = 3.3, height = 6, dpi=300)

```

#### b. Salinity
```{r}
df_raw <- VT3  %>%
  select (Date, Depth, Salinity) %>%
  rename (Variable = Salinity) %>%
  drop_na()

# Interpolate using GAM with tensor product smooth (te)
mod1 <- gam(Variable ~ te(as.numeric(Date), Depth), data = df_raw)
mod2 <- gam(Variable ~ te(as.numeric(Date), Depth, k = 20), data = df_raw)

# Make data for predicting values (every 1 day and every 1 meter)
df_pred <- expand.grid(
  list(
    Date = seq(min(df_raw$Date), max(df_raw$Date), by = "days"),
    Depth = seq(min(df_raw$Depth), max(df_raw$Depth))))

# Use models to predict
df_pred$Var_gam1 <- predict(mod1, df_pred)
df_pred$Var_gam2 <- predict(mod2, df_pred)

# gam.check(mod2)
plot.gam (mod2)

# Upper 30m only
raw_30 <- df_raw %>%
  filter (Depth >= -30) 
pred_30 <- df_pred %>%
  filter (Depth >= -30) 

# whole water column
Salt_VT3 <- ggplot(df_pred, aes(x = Date, y = Depth)) + 
  geom_raster(aes(fill = Var_gam2), alpha = 0.5) +            # interpolation
    scale_fill_gradientn(colours = jet.colors(20), limits = c(9, 36)) + 
  geom_point(data = df_raw, aes(colour = Variable), shape = "-", size = rel(6)) +  # raw
    scale_colour_gradientn(colours = jet.colors(20), limits = c(0, 36), guide = 'none') +
  geom_contour(data = df_pred, aes(z = Var_gam2), color = "white", alpha = 0.5) +
      labs(title = "VT3", x = "", y = "Salinity (PSU)", fill = "") +
      scale_x_date(date_breaks = "2 month", date_labels =  "%b %y") +
      theme(axis.text.x=element_text(angle=60, hjust=1), plot.title = element_text(hjust = 0.5))

plot (Salt_VT3)
#?ggsave
# ggsave ("Figures/_Nov19/Contour_VT3_Salt.pdf", dpi=300)

# upper 30m
Salt_VT3_30 <- ggplot(pred_30, aes(x = Date, y = Depth)) + 
  geom_raster(aes(fill = Var_gam2), alpha = 0.7) +            # interpolation
    scale_fill_gradientn(colours = jet.colors(20), limits = c(9,36)) + 
  geom_point(data = raw_30, aes(colour = Variable), shape = "-", size = rel(6)) +  # raw
    scale_colour_gradientn(colours = jet.colors(20), limits = c(0, 36), guide = 'none') +
  geom_contour(data = pred_30, aes(z = Var_gam2), color = "white", alpha = 0.5) +
      labs(title = "VT3", x = "", y = "Salinity (PSU)", fill = "") +
      scale_x_date(date_breaks = "2 month", date_labels =  "%b %y") +
      theme(axis.text.x=element_text(angle=60, hjust=1), plot.title = element_text(hjust = 0.5))

plot (Salt_VT3_30)
# ggsave ("Figures/_Nov19/Contour_VT3_Salt_30m.pdf", width = 3.3, height = 6, dpi=300)

```
#### c. Oxygen
```{r}

df_raw <- VT3  %>%
  select (Date, Depth, O2_sat) %>%
  rename (Variable = O2_sat) %>%
  drop_na()

# Interpolate using GAM with tensor product smooth (te)
mod1 <- gam(Variable ~ te(as.numeric(Date), Depth), data = df_raw)
mod2 <- gam(Variable ~ te(as.numeric(Date), Depth, k = 20), data = df_raw)

# Make data for predicting values (every 1 day and every 1 meter)
df_pred <- expand.grid(
  list(
    Date = seq(min(df_raw$Date), max(df_raw$Date), by = "days"),
    Depth = seq(min(df_raw$Depth), max(df_raw$Depth))))

# Use models to predict
df_pred$Var_gam1 <- predict(mod1, df_pred)
df_pred$Var_gam2 <- predict(mod2, df_pred)

# Upper 30m only
raw_30 <- df_raw %>%
  filter (Depth >= -30) 
pred_30 <- df_pred %>%
  filter (Depth >= -30) 

range (df_raw$Variable)

# whole water column
O2_VT3 <- ggplot(df_pred, aes(x = Date, y = Depth)) + 
  geom_raster(aes(fill = Var_gam2), alpha = 0.5) +            # interpolation
    scale_fill_gradientn(colours = jet.colors(20), limits = c(15, 122)) + 
  geom_point(data = df_raw, aes(colour = Variable), shape = "-", size = rel(6)) +  # raw
    scale_colour_gradientn(colours = jet.colors(20), limits = c(15, 122), guide = 'none') +
  geom_contour(data = df_pred, aes(z = Var_gam2), color = "white", alpha = 0.5) +
      labs(title = "VT3", x = "", y = "O2 (ml/L)", fill = "") +
      scale_x_date(date_breaks = "2 month", date_labels =  "%b %y") +
      theme(axis.text.x=element_text(angle=60, hjust=1), plot.title = element_text(hjust = 0.5))

plot (O2_VT3)
#?ggsave
# ggsave ("Figures/_Nov19/Contour_VT3_O2.pdf", dpi=300)

# upper 30m
O2_VT3_30 <- ggplot(pred_30, aes(x = Date, y = Depth)) + 
  geom_raster(aes(fill = Var_gam2), alpha = 0.7) +            # interpolation
    scale_fill_gradientn(colours = jet.colors(20), limits = c(15, 122)) + 
  geom_point(data = raw_30, aes(colour = Variable), shape = "-", size = rel(6)) +  # raw
    scale_colour_gradientn(colours = jet.colors(20), limits = c(15, 122), guide = 'none') +
  geom_contour(data = pred_30, aes(z = Var_gam2), color = "white", alpha = 0.5) +
      labs(title = "VT3", x = "", y = "O2 (ml/L)", fill = "") +
      scale_x_date(date_breaks = "2 month", date_labels =  "%b %y") +
      theme(axis.text.x=element_text(angle=60, hjust=1), plot.title = element_text(hjust = 0.5))

plot (O2_VT3_30)
# ggsave ("Figures/_Nov19/Contour_VT3_O2_30m.pdf", width = 3.3, height = 6, dpi=300)



```

#### d. DOC 
```{r}
#check range
range (df_m$DOC, na.rm = T)

# make dataset for interpolation and plotting
df_raw <- VT3  %>%
  select (Date, Depth, DOC) %>%
  rename (Variable = DOC) %>%
  drop_na()

# Interpolate using GAM with tensor product smooth (te)
mod1 <- gam(Variable ~ te(as.numeric(Date), Depth), data = df_raw)
mod2 <- gam(Variable ~ te(as.numeric(Date), Depth, k = 6), data = df_raw)

#?choose.k

# gam.check(mod2)
plot.gam (mod2) # does not look nice

# Make data for predicting values (every 1 day and every 1 meter)
df_pred <- expand.grid(
  list(
    Date = seq(min(df_raw$Date), max(df_raw$Date), by = "days"),
    Depth = seq(min(df_raw$Depth), max(df_raw$Depth))))

# Use models to predict
df_pred$Var_gam1 <- predict(mod1, df_pred)
df_pred$Var_gam2 <- predict(mod2, df_pred)

# whole water column
DOC_VT3 <- ggplot(df_pred, aes(x = Date, y = Depth)) + 
  geom_raster(aes(fill = Var_gam2), alpha = 0.5) +            # interpolation
    scale_fill_gradientn(colours = jet.colors(20), limits = c(0.7, 3.4)) + 
  geom_point(data = df_raw, aes(colour = Variable), shape = "-", size = rel(6)) +  # raw
    scale_colour_gradientn(colours = jet.colors(20), limits = c(0.7, 3.4), guide = 'none') +
  geom_contour(data = df_pred, aes(z = Var_gam2), color = "white", alpha = 0.5) +
      labs(title = "VT3", x = "", y = "DOC (mg/L)", fill = "") +
      scale_x_date(date_breaks = "2 month", date_labels =  "%b %y") +
      theme(axis.text.x=element_text(angle=60, hjust=1), plot.title = element_text(hjust = 0.5))

plot (DOC_VT3)

# the interpolated values does capture the variabilit in the measured values. But not possible to increase K further than 6.5 because of insufficient unique values... is there any fix, or need to use different interpolation????


#?ggsave
#ggsave ("Figures/_Nov19/Contour_VT3_DOC.pdf", dpi=300)

```

#### d2. Base function interpol
```{r}

all.dates <- seq(from = min(df_raw$Date), to = max(df_raw$Date), by = "day")
all.depths <- -seq(0, 30, 1)

interpol <- with(df_raw,
                 interp(x = Date, y = Depth, z = Variable, 
                        xo = all.dates, yo = all.depths)
)

# Plot 1
# Test plot using image.plot
# Need to turn around the results so y values are in increasing order (because
# that is what image() likes)
interpol$y <- interpol$y[seq(length(interpol$y), 1)]
interpol$z <- interpol$z[, seq(length(interpol$y), 1)]   
fields::image.plot(interpol)

# xtabs(~Date + Depth, df_raw)

pick_date <- unique(df_raw$Date)[20]
sel_x <- interpol$x == pick_date
plot(interpol$y, interpol$z[sel_x,], type = "l", 
     xlab = "Depth", ylab = "Interpolation", 
     main = as.Date(interpol$x[sel_x], origin = "1970-01-01"),
     ylim = range(df_raw$Variable)
     )
points(Variable ~ Depth, subset(df_raw, Date == pick_date))

## ggplot style
# df_interpol
df <- as.data.frame(interpol$z)
colnames(df) <- interpol$y
df$Date <- interpol$x

df_pred <- df %>%
  tidyr::gather("Depth", "Var_int", -Date)

# Restore our variables to proper format....
df_pred$Depth <- as.numeric(df_pred$Depth)
df_pred$Date <- as.Date(df_pred$Date, origin = "1970-01-01")

# code from above, except that "Var_gam2" is replaced by "Var_int"
DOC_VT3 <- ggplot(df_pred, aes(x = Date, y = Depth)) + 
  geom_raster(aes(fill = Var_int), alpha = 0.5) +            # interpolation
    scale_fill_gradientn(colours = jet.colors(20), limits = c(0.7, 3.4)) + 
  geom_point(data = df_raw, aes(colour = Variable), shape = "-", size = rel(6)) +  # raw
    scale_colour_gradientn(colours = jet.colors(20), limits = c(0.7, 3.4), guide = 'none') +
  geom_contour(data = df_pred, aes(z = Var_int), color = "white", alpha = 0.5) +
      labs(title = "VT3", x = "", y = "DOC (mg/L)", fill = "") +
      scale_x_date(date_breaks = "2 month", date_labels =  "%b %y") +
      theme(axis.text.x=element_text(angle=60, hjust=1), plot.title = element_text(hjust = 0.5))

plot(DOC_VT3) 


```



#### d3. interp() from package akima  
* Problem: interp()is not meant for regular (rectangular) gridded data! (source: ?interp)  
* Quite big artefacts created  
```{r}

library(akima)

all.dates <- seq(from = min(df_raw$Date), to = max(df_raw$Date), by = "day")
all.depths <- -seq(0, 30, 1)

interpol <- with(df_raw,
                 interp(x = Date, y = Depth, z = Variable, 
                        xo = all.dates, yo = all.depths,
                        linear = FALSE, jitter = 0.5, jitter.iter = 20)
                 )

# Test plot using image
# Need to turn around the results so y values are in increasing order (because
# that is what image() likes)
interpol$y <- interpol$y[seq(length(interpol$y), 1)]
interpol$z <- interpol$z[, seq(length(interpol$y), 1)]   
image(interpol)

## ggplot style
# df_interpol
df <- as.data.frame(interpol$z)
colnames(df) <- interpol$y
df$Date <- interpol$x

df_pred <- df %>%
  tidyr::gather("Depth", "Var_int", -Date)


# Restore our variables to proper format....
df_pred$Depth <- as.numeric(df_pred$Depth)
df_pred$Date <- as.Date(df_pred$Date, origin = "1970-01-01")

# code from above, except that "Var_gam2" is replaced by "Var_int"
DOC_VT3 <- ggplot(df_pred, aes(x = Date, y = Depth)) + 
  geom_raster(aes(fill = Var_int), alpha = 0.5) +            # interpolation
    scale_fill_gradientn(colours = jet.colors(20), limits = c(0.7, 3.4)) + 
  geom_point(data = df_raw, aes(colour = Variable), shape = "-", size = rel(6)) +  # raw
    scale_colour_gradientn(colours = jet.colors(20), limits = c(0.7, 3.4), guide = 'none') +
  geom_contour(data = df_pred, aes(z = Var_int), color = "white", alpha = 0.5) +
      labs(title = "VT3", x = "", y = "DOC (mg/L)", fill = "") +
      scale_x_date(date_breaks = "2 month", date_labels =  "%b %y") +
      theme(axis.text.x=element_text(angle=60, hjust=1), plot.title = element_text(hjust = 0.5))

plot(DOC_VT3) 

```


#### d4. interp.surface() from package fields  
* Also tried bicubic.grid() from package akima but it fails if there are NAs in data
* interp.surface() works but makes a hole where 30 m data is lacking  
* interp.surface() also ran with higher resolution (0.25 days * 0.25 meters) but that doesn't really help  

```{r}
df_raw_matrix <- df_raw %>%
  filter(!Depth %in% -2) %>%
  select(Date, Depth, Variable) %>%
  spread(Depth, Variable)

#
# tried bicubic.grid() from package akima 
#
if (FALSE){
  interpol2 <- 
    akima::bicubic.grid(
      x = df_raw_matrix$Date %>% as.vector(),
      y = colnames(df_raw_matrix %>% select(-Date) %>% as.vector()),
      z = as.matrix(df_raw_matrix %>% select(-Date)),
      nx = 40,
      ny = 40)
  # RESULT:
  # Error in bicubic(x, y, z, xy[, 1], xy[, 2]) : 
  #   NA/NaN/Inf in foreign function call (arg 6)
}

df_for_interpol <- expand.grid(Date = all.dates, Depth = all.depths)

interpol3 <- 
  fields::interp.surface(
    list(
      x = df_raw_matrix$Date %>% as.vector(),
      y = colnames(df_raw_matrix %>% select(-Date) %>% as.vector()),
      z = as.matrix(df_raw_matrix %>% select(-Date))
    ),
    df_for_interpol
  )

df_pred <- cbind(df_for_interpol, Var_int = interpol3)


# code from above, except that "Var_gam2" is replaced by "Var_int"
DOC_VT3 <- ggplot(df_pred, aes(x = Date, y = Depth)) + 
  geom_raster(aes(fill = Var_int), alpha = 0.5) +            # interpolation
    scale_fill_gradientn(colours = jet.colors(20), limits = c(0.7, 3.4)) + 
  geom_point(data = df_raw, aes(colour = Variable), shape = "-", size = rel(6)) +  # raw
    scale_colour_gradientn(colours = jet.colors(20), limits = c(0.7, 3.4), guide = 'none') +
  geom_contour(data = df_pred, aes(z = Var_int), color = "white", alpha = 0.5) +
      labs(title = "VT3", x = "", y = "DOC (mg/L)", fill = "") +
      scale_x_date(date_breaks = "2 month", date_labels =  "%b %y") +
      theme(axis.text.x=element_text(angle=60, hjust=1), plot.title = element_text(hjust = 0.5))

plot(DOC_VT3) 


#
# Higher resolution
#

all.dates2 <- seq(from = min(as.numeric(df_raw$Date)), 
                  to = max(as.numeric(df_raw$Date)), 
                  by = 0.25)
all.depths2 <- -seq(0, 30, 0.25)

df_for_interpol <- expand.grid(Date = all.dates2, Depth = all.depths2)

interpol3 <- 
  fields::interp.surface(
    list(
      x = df_raw_matrix$Date %>% as.vector(),
      y = colnames(df_raw_matrix %>% select(-Date) %>% as.vector()),
      z = as.matrix(df_raw_matrix %>% select(-Date))
    ),
    df_for_interpol
  )

df_pred <- cbind(df_for_interpol, Var_int = interpol3)

# Date on time format (POSIXct = seconds since 1st Jan 1970)
df_pred$Date <- as.POSIXct(60*60*24*df_pred$Date, origin = "1970-01-01", tz = "GMT")
# str(df_pred)

# code from above (part 'd'), except that 
#    "Var_gam2" is replaced by "Var_int"
#    scale_x_date replaced by scale_x_datetime
#    df_raw 'Date' variable  changed from "Date" format to POSIXct (see in aes() specification)
DOC_VT3 <- ggplot(df_pred, aes(x = Date, y = Depth)) + 
  geom_raster(aes(fill = Var_int), alpha = 0.5) +            # interpolation
    scale_fill_gradientn(colours = jet.colors(20), limits = c(0.7, 3.4)) + 
  geom_point(data = df_raw, aes(x = as.POSIXct(Date), colour = Variable), 
             shape = "-", size = rel(6)) +  # raw
      scale_colour_gradientn(colours = jet.colors(20), limits = c(0.7, 3.4), guide = 'none') +
  geom_contour(data = df_pred, aes(z = Var_int), color = "white", alpha = 0.5) +
      labs(title = "VT3", x = "", y = "DOC (mg/L)", fill = "") +
      scale_x_datetime(date_breaks = "2 month", date_labels =  "%b %y") +
      theme(axis.text.x=element_text(angle=60, hjust=1), plot.title = element_text(hjust = 0.5))

plot(DOC_VT3) 

```




#### e. Chla 
```{r}

#check range
#range (df_m$DOC, na.rm = T)

# make dataset for interpolation and plotting
df_raw <- VT3  %>%
  select (Date, Depth, Chla) %>%
  rename (Variable = Chla) %>%
  drop_na()

# Interpolate using GAM with tensor product smooth (te)
mod1 <- gam(Variable ~ te(as.numeric(Date), Depth), data = df_raw)
mod2 <- gam(Variable ~ te(as.numeric(Date), Depth, k = 5), data = df_raw)

#?choose.k
#gam.check(mod2)
#plot.gam (mod2) # does not look nice

# Make data for predicting values (every 1 day and every 1 meter)
df_pred <- expand.grid(
  list(
    Date = seq(min(df_raw$Date), max(df_raw$Date), by = "days"),
    Depth = seq(min(df_raw$Depth), max(df_raw$Depth))))

# Use models to predict
df_pred$Var_gam1 <- predict(mod1, df_pred)
df_pred$Var_gam2 <- predict(mod2, df_pred)

# whole water column
Chla_VT3 <- ggplot(df_pred, aes(x = Date, y = Depth)) + 
  geom_raster(aes(fill = Var_gam2), alpha = 0.5) +            # interpolation
    scale_fill_gradientn(colours = jet.colors(20)) + 
  geom_point(data = df_raw, aes(colour = Variable), shape = "-", size = rel(6)) +  # raw
    scale_colour_gradientn(colours = jet.colors(20), guide = 'none') +
  geom_contour(data = df_pred, aes(z = Var_gam2), color = "white", alpha = 0.5) +
      labs(title = "VT3", x = "", y = "Chla (ug/L)", fill = "") +
      scale_x_date(date_breaks = "2 month", date_labels =  "%b %y") +
      theme(axis.text.x=element_text(angle=60, hjust=1), plot.title = element_text(hjust = 0.5))

plot (Chla_VT3)
#even worse...

#ggsave ("Figures/_Oct19/Contour_VT3_Chla.png", width = 8, height = 6, dpi=500)
```




