---
title: "11_Vannmiljo_2019"
author: "DHJ"
date: "16 3 2020"
output: html_document
---

## 1. Libraries etc.
```{r}
library(readxl)
library(dplyr)
library(ggplot2)
library(lubridate)
library(sp)        # SpatialPoints(), CRS(), spTransform()

crs_longlat <- "+proj=longlat"
crs_utm <- "+proj=utm +zone=33 +ellps=WGS84 +datum=WGS84 +units=m"

```


## 2. Test files  
Exported from Vannmiljø 16.03.2020  
- Medium = "Saltvann"
- Øvre dyp = nedre dyp = between 4 and 11
- Parameter = "salinitet" (dat1) and "nitrat" + "nitrat + nitritt" (dat2)   

### Data
```{r}
dat1 <- read_excel("Input_data/VannmiljoEksport_2019test_salinity.xlsx")
dat2 <- read_excel("Input_data/VannmiljoEksport_2019test_nitrate.xlsx")
dat <- bind_rows(dat1, dat2)

# Seems that dyp has been set to NA where Dyp is really zero
dat$Ovre_dyp[is.na(dat$Ovre_dyp)] <- 0
dat$Nedre_dyp[is.na(dat$Nedre_dyp)] <- 0

# Convert Verdi (value) to numbers (first have to change decimal comma to 'full stop')
x <- as.numeric(sub(",", ".", dat$Verdi, fixed = TRUE))
sum(is.na(x))  # 0, so all ar ok
dat$Verdi <- x

x <- as.numeric(sub(",", ".", dat$Ovre_dyp, fixed = TRUE))
sum(is.na(x))  # 0, so all ar ok
dat$Ovre_dyp <- x

x <- as.numeric(sub(",", ".", dat$Nedre_dyp, fixed = TRUE))
sum(is.na(x))  # 0, so all ar ok
dat$Nedre_dyp <- x

# Make time variable
dat$Time <- ymd_hms(dat$Tid_provetak)

# Fix Siktedyp depth to zero
sel <- dat$Parameter_navn == "Siktedyp"; sum(sel); # table(dat$Nedre_dyp[sel])
dat$Nedre_dyp[sel] <- 0

#
# Add ordinary long/lat (transform from UTM zone 33 to Long,Lat)
#
SP.utm <- SpatialPoints(dat[,c("UTM33 Ost (X)", "UTM33 Nord (Y)")], 
                           proj4string=CRS(crs_utm)
                           )
SP.longlat <- spTransform(SP.utm, CRS(crs_longlat))
dat$Long <- SP.longlat@coords[,1]
dat$Lat <- SP.longlat@coords[,2
                             
```
### Table  
```{r}

dat %>%
  xtabs(~Parameter_navn, .)

dat %>%
  filter(Long > 7.5 & Lat < 60) %>%
  xtabs(~Parameter_navn + Vannlokalitet, .)

dat %>%
  filter(Long > 7.5 & Lat < 60) %>%
  xtabs(~Oppdragstaker + Vannlokalitet, .)

```

### Map
```{r}

map_df <- dat %>% 
  count(Vannlokalitet, Long, Lat) %>% 
  mutate(Popup = paste0(Vannlokalitet, "<br>Long,lat = ", round(Long, 4), ", ", round(Lat, 4), ""))

#
# Test plot
#
library(leaflet)
leaflet() %>%
  addTiles() %>%  # Default OpenStreetMap map tiles
  addMarkers(lng = map_df$Long, lat = map_df$Lat,
             popup = map_df$Popup)

map_df_sel <- map_df %>%
  filter(Long > 7.5 & Lat < 60)

leaflet() %>%
  addTiles() %>%  # Default OpenStreetMap map tiles
  addMarkers(lng = map_df_sel$Long, lat = map_df_sel$Lat,
             popup = map_df_sel$Popup)
```

### Copy codes to Vannmiljø  
Used this to copy-paste water locality codes to Vannmiljø   
- it didn't work, though.... (some error message in Vannmiljø)  
- so I ended up making a search for all medium = Saltvann for 2019 for each Oppdragstaker    
```{r}

# writeLines(unique(dat$Vannlokalitet_kode), "clipboard")


```


## 3. 2019 files  
Exported from Vannmiljø 16.03.2020  
- Medium = "Saltvann"
- Oppdragstaker = COWI, Dnv or NIRAS (see test files)
### Data
```{r}
dat1 <- read_excel("Input_data/VannmiljoEksport_2019_COWI.xlsx")
dat2 <- read_excel("Input_data/VannmiljoEksport_2019_DnV.xlsx")
dat3 <- read_excel("Input_data/VannmiljoEksport_2019_NIRAS.xlsx")
dat <- bind_rows(dat1, dat2, dat3)

# Seems that dyp has been set to NA where Dyp is really zero
dat$Ovre_dyp[is.na(dat$Ovre_dyp)] <- 0
dat$Nedre_dyp[is.na(dat$Nedre_dyp)] <- 0

# Convert Verdi (value) to numbers (first have to change decimal comma to 'full stop')
x <- as.numeric(sub(",", ".", dat$Verdi, fixed = TRUE))
sum(is.na(x))  # 0, so all ar ok
dat$Verdi <- x

x <- as.numeric(sub(",", ".", dat$Ovre_dyp, fixed = TRUE))
sum(is.na(x))  # 0, so all ar ok
dat$Ovre_dyp <- x

x <- as.numeric(sub(",", ".", dat$Nedre_dyp, fixed = TRUE))
sum(is.na(x))  # 0, so all ar ok
dat$Nedre_dyp <- x

# Make time variable
dat$Time <- ymd_hms(dat$Tid_provetak)

# Fix Siktedyp depth to zero
sel <- dat$Parameter_navn == "Siktedyp"; sum(sel); # table(dat$Nedre_dyp[sel])
dat$Nedre_dyp[sel] <- 0

#
# Add ordinary long/lat (transform from UTM zone 33 to Long,Lat)
#
SP.utm <- SpatialPoints(dat[,c("UTM33 Ost (X)", "UTM33 Nord (Y)")], 
                           proj4string=CRS(crs_utm)
                           )
SP.longlat <- spTransform(SP.utm, CRS(crs_longlat))
dat$Long <- SP.longlat@coords[,1]
dat$Lat <- SP.longlat@coords[,2]
                             
```
### Table  
```{r}

dat %>%
  xtabs(~Parameter_navn, .)

dat %>%
  filter(Long > 7.5 & Lat < 60) %>%
  xtabs(~Parameter_navn + Vannlokalitet, .)

dat %>%
  filter(Long > 7.5 & Lat < 60) %>%
  xtabs(~Oppdragstaker + Vannlokalitet, .)

```

### Map
```{r}

map_df <- dat %>% 
  count(Vannlokalitet, Long, Lat) %>% 
  mutate(Popup = paste0(Vannlokalitet, "<br>Long,lat = ", round(Long, 4), ", ", round(Lat, 4), ""))

#
# Test plot
#
library(leaflet)
leaflet() %>%
  addTiles() %>%  # Default OpenStreetMap map tiles
  addMarkers(lng = map_df$Long, lat = map_df$Lat,
             popup = map_df$Popup)

map_df_sel <- map_df %>%
  filter(Long > 7.5 & Lat < 60)

leaflet() %>%
  addTiles() %>%  # Default OpenStreetMap map tiles
  addMarkers(lng = map_df_sel$Long, lat = map_df_sel$Lat,
             popup = map_df_sel$Popup)

```

## 4. Save  
```{r}

openxlsx::write.xlsx(dat, "Data/11_Vannmiljo_2019.xlsx")

```

