---
title: "Test read ocean color NetCDF"
author: "DHJ"
date: "16 8 2019"
output: 
  html_document:
    keep_md: true
---


## Library
```{r}
library(ncdf4)
library(maps)
library(mapdata)
library(dplyr)
library(ggplot2)
```

```{r}
filepath = 'http://thredds.met.no/thredds/dodsC/metusers/arildb/MARTINI800_prov_v2.ncml' # The OPENDAP URL
nc <- nc_open(filepath)

```

```{r}
Sys.getenv("JAVA_HOME")
```


## 8-day data  
Note that the 8-day means have time given as UNIX time (seconds sonce 01-01-1970), while the daily means have time given as days since 01-01-1900 (i.e. Excel time)  

### Pick file  
8-day means L4 Chl a concentration, saved in 3 files. Note that one obs. time, the date 2001-11-18, is not included (as the NetCDF was corrupt). So there is a one-observation gap between files 1 and 2.
```{r}
datafolder <- "Input_data/Copernicus"
fns <- dir(datafolder, ".nc")
fns
fn <- "dataset-oc-atl-chl-multi_cci-l4-chl_1km_8days-rep-v02 1997 - 9nov2001 .nc"
```

### Explore file
```{r}
# ?ncdf4::ncvar_get

nc <- nc_open(paste(datafolder, fn, sep = "/"))
nc$nvars
str(nc$var, 1)
str(nc$dim, 1)
str(nc$dim, 2)

# plot(nc$dim$lat$vals)
# plot(nc$dim$lon$vals)
# plot(nc$dim$time$vals)
```

### Read data
```{r}
x <- ncvar_get(nc, "CHL")  # reads all 
dim(x)

# Range of data
zlim <- range(x, na.rm = TRUE)
zlim_plot = log10(zlim)

# Maximum is observed where and when?
arrayInd(which.max(x), dim(x))

# Close connection to nc file
nc_close(nc)
```

### Plot map for a given day
```{r}
# Time step number
i <- 32

# Time in R format 
time <- as.POSIXct(nc$dim$time$vals, origin = "1970-01-01", tz = "GMT")
time

# Data for this day
# The '[,seq(dim(x)[2],1)]' part flips the matrix so it displys with North upwards
zvar <- x[,,i][,seq(dim(x)[2],1)]

x_day <- list(
  x = nc$dim$lon$vals,
  y = rev(nc$dim$lat$vals),
  z = log10(zvar)
)

fields::image.plot(x_day, main = time[i], zlim = zlim_plot)
map("worldHires", regions = c("Norway", "Sweden"), add = TRUE)
```
### Plot mean + 20/80 percentiles
```{r}
df_stat <- data.frame(
  time = as.POSIXct(nc$dim$time$vals, origin = "1970-01-01", tz = "GMT"),
  mean = 10^apply(log10(x), 3, mean, na.rm = TRUE),
  p20 = apply(x, 3, quantile, probs = 0.2, na.rm = TRUE),
  p80 = apply(x, 3, quantile, probs = 0.8, na.rm = TRUE)
  )
plot(p80 ~ time , type = "l", col = "grey50", log = "y", data = df_stat, 
     main = "Chl a", ylab = "Mean / Perc20 / Perc80")
lines(mean ~ time, type = "o", pch = 4, data = df_stat)
lines(p20 ~ time, col = "grey50", data = df_stat)
# length(time)


```


## Daily data  
One file per year (downloaded using Python, using the batch file 'copernicus_download_daily_chla_skagerrak.bat')  
Get data  
```{r}
datafolder <- "Input_data/Copernicus/Daily"
fns <- dir(datafolder, ".nc")
# fns
fn <- "dataset-oc-atl-chl-multi-l4-oi_1km_daily-rep-v02_2017.nc"

nc <- nc_open(paste(datafolder, fn, sep = "/"))

x <- ncvar_get(nc, "CHL")  # reads all 
dim(x)

# Range of data
zlim <- range(x, na.rm = TRUE)
zlim_plot = log10(zlim)

t <- nc$dim$time$vals
range(t)

# Maximum is observed where and when?
arrayInd(which.max(x), dim(x))

# Close connection to nc file
nc_close(nc)

# Get mean + 20/80 percentiles
df_stat <- data.frame(
  time = as.POSIXct(nc$dim$time$vals*24*60*60, origin = "1900-01-01", tz = "GMT"),
  mean = 10^apply(log10(x), 3, mean, na.rm = TRUE),
  p10 = apply(x, 3, quantile, probs = 0.1, na.rm = TRUE),
  p20 = apply(x, 3, quantile, probs = 0.2, na.rm = TRUE),
  p80 = apply(x, 3, quantile, probs = 0.8, na.rm = TRUE),
  p90 = apply(x, 3, quantile, probs = 0.9, na.rm = TRUE)
  )

```

```{r}
ggplot(df_stat, aes(x = time)) + 
  geom_ribbon(aes(ymin = p10, ymax = p90), fill = "grey70") +
  geom_ribbon(aes(ymin = p20, ymax = p80), fill = "grey50") +
  geom_line(aes(y = mean)) +
  coord_cartesian(ylim = c(0, 20))
```


