---
title: "Workshop - reading thredds data using Python"
author: "DHJ"
date: "23 9 2019"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true

---

# Reading from met.no's thredds server using Python chunks  

Results from met.no's oceanographic models are stored as huge NetCDF files in the `ncml` format. Those files are too big for downloading; instead one reads just the part of the data that you are interested in. The usual R library for the NetCDF format (`ncdf4`) can read `.nc` files, but not the `ncml` format. A new R libarry called `LoadeR` (that I know of) _can_ read `ncml` files, but was found to not handle the particular format used in the oceanographic model results.   
  
However, Python has good tools for this. Fortunately, calling Python from R is pretty easy, and has gotten very elegant using the reticulate package. In this markdown document we use code chunks of two types, R code chunks and Python code chunks.         

The Python code inside the Python chunks are based on Python examples from a Jupyter notebook by  
Ann-Kristin Sperrevik, see 
* https://nbviewer.jupyter.org/github/annks/Examples/blob/master/MartiniControlRun.ipynb    
and   
* https://github.com/annks/Examples  
See Ann-Kristin's notebook for more explanation of the Python parts.    



## 1. R part

### a. Libraries
```{r, results='hide', message=FALSE}
library(reticulate)
use_python("C:/WinPython/WPy64-3720/python-3.7.2.amd64/python.exe")   # my Python installation

library(readxl)
library(ggplot2)
library(lubridate)
library(dplyr)
library(purrr)


```


### b. Get raw data and locations
* Read hard-bottom data, `data_hardbottom`  
* Then we extract location of the most used stations, `data_loc`  
    - We use count() to get one line per station  
    - count() automatocally adds the variable n for sample size  
    - We then use n to get only stations with >= 500 lines of data  
    
```{r}
data_hardbottom <- read_excel("Test_data/data_hardbottom.xlsx")

data_stations <- data_hardbottom %>%
  count(STATION_CODE, STATION_NAME, LONGITUDE, LATITUDE) %>%
  filter(n > 500)
```

### c. Plot stations in leaflet
```{r}
library(leaflet)

leaflet() %>%
  addTiles() %>%
  addMarkers(lng = data_stations$LONGITUDE, lat = data_stations$LATITUDE,
             popup = paste(data_stations$STATION_CODE, data_stations$STATION_NAME))

```

### d. Select one location     
Færder fyr  
* Note: this returns a one-row dat frame
```{r}
selected_location <- data_stations %>%
  filter(STATION_CODE == "HT4")

selected_location
```

## 2. Python part

* Python code is put inside code chunks marked 'python' instead of 'R'
* Note that you have to avoid Norwegian letters in the Python code, even in the comments - it causes errors (there is probably a way to fix this)  

### a. Connect to thredds server  
```{python}
import numpy as np     # Package for scientific computing 
import pandas as pd    # Package for data frames 
import matplotlib.pyplot as plt   
from datetime import datetime,timedelta

from netCDF4 import Dataset #  This is handy for working with netCDF files

import cartopy.crs as ccrs
import cartopy.feature as cfeature

import roppy

filepath = 'http://thredds.met.no/thredds/dodsC/metusers/arildb/MARTINI800_prov_v2.ncml' # The OPENDAP URL

filehandle = Dataset(filepath) # open for reading 
grid = roppy.SGrid(filehandle) # Create a grid object for our file

```

### b. List all variables  
Lots to choose from!  
```{python}
print(filehandle.variables.keys())

```

### c. Get info on one variable
```{python}
filehandle.variables['P3_Chl']
```

### d. Get all variables' long names  
Note: commented out, as it generates a very long list  
```{python, result = 'hide'}
keys = filehandle.variables.keys()
print('Number of variables:', len(keys))
print()

# for key in keys:
#   print(key, '=', filehandle.variables[key].long_name)

```



### e. Access longitude, latitude and time  and dimensions    
* This doesn't actually download the data, it just creates a 'pointer' to them  
```{python}
# Access to the longitude, latitude coordinates

lon = filehandle.variables['lon_rho']
lat = filehandle.variables['lat_rho']
ocean_time = filehandle.variables['ocean_time']

```

### f. Explore dimensions 
```{python}
# Let's see what the dimensions of this variable is by looping over the dimensions
print('Time dimensions:')
for dimension in ocean_time.dimensions:
    # We'll print both dimension name, and the size of th dimension
    print('{}: {}'.format( dimension,  len(filehandle.dimensions[dimension] ) ) )

# And lets look at longitude as well:
print('\nLongitude (or latitude) dimensions:')   # \n is just "line shift"
for dimension in lon.dimensions:
    print( '{}: {}'.format( dimension,  len( filehandle.dimensions[dimension] ) ) )

# And finally let's look at one 'ordinary' variable:
print('\nDimensions of the temperature variable:')
temp = filehandle.variables['temp']
for dimension in temp.dimensions:
    print( '{}: {}'.format( dimension,  len( filehandle.dimensions[dimension] ) ) )

```



### g. Find grid cell closest to our chosen station in R      
* Inside a Python chunk, we can access an R Object (values, vectors, dataframes) by putting `r.` in front of the objects name  
    - E.g., we access R object `selected_location` by `r.selected_location`  
* R data frames are converted to a Pandas DataFrame in Python   
* In a Pandas DataFrame, you access a value by ['column_name'][row_number]  
* Here we will access the R data frame `selected_location` (which has only one line)
* We need to access longitude number 1 and latitude number 1 of this data frame -
but be aware that Python starts counting at zero
* So to get the first Longitude value, we write  
    - R version 1: selected_location[1, 'LONGITUDE'] - OR  
    - R version 2: selected_location$LONGITUDE[1]  
    - Python: r.selected_location['LONGITUDE'][0]  

```{python}
# The decimal degree location of the selected station:
selected_lon = r.selected_location['LONGITUDE'][0]
selected_lat = r.selected_location['LATITUDE'][0]

# We need to find the grid location closest to the selected coordinates.
# For the sake of simplicity, let's calculate the sum of the absolute 
# differences between all grid points latitude and longitude and the 
# selected coordinates. 

position_diff = np.abs( lat[:] - selected_lat ) + np.abs( lon[:] - selected_lon )

# This line will find the indices of the minimum value in 
i, j = np.unravel_index( position_diff.argmin(), position_diff.shape )

print('Grid indices of grid point closest to selected location: {}, {}\n'.format(i, j))
print('Grid point longitude: {}'.format(lon[i,j]))
print('Grid point latitude: {}'.format(lat[i,j]))

```


### h. Getting surface values    
```{python}
# The dimensions are in (t,z,y,x) order and as indexi,g in Python starts from zero, 
# axis = 1 points to the z-dimension. 

times = np.arange(0,360,30) 

# temp = potential temperaturen
# salt = salinity
# N1_p = phosphate/phosphorus
# N3_n = nitrate/nitrogen
# chla = chlorophyll A
data_loc = pd.DataFrame({
  'Station': r.selected_location['STATION_CODE'][0],  
  'i': i,
  'j': j,
  'Time_num': filehandle.variables['ocean_time'][times],
  'Temp': filehandle.variables['temp'][times,0,i,j],
  'Salinity': filehandle.variables['salt'][times,0,i,j],
  'PO4': filehandle.variables['N1_p'][times,0,i,j],
  'NO3': filehandle.variables['N3_n'][times,0,i,j]
  })

```

## 3. Bring data back to R  

### a. Collect the data from Python  
```{r}
py$data_loc

```

### b. Fix time variable  
As explained in Ann Kristin's notebook, time is given in seconds since 1. Jan. 1970  
```{r}
data_loc <- py$data_loc
data_loc$Date <- as.POSIXct(data_loc$Time_num, origin = "1970-01-01", tz = "UTC")
```

### c. A couple of plots
```{r}
plot(Temp ~ Date, type = "b", data_loc)
plot(Temp ~ Salinity, type = "b", data_loc)
```

## 4. Get values by calling Python function   

### a. Get the function in file "32_Function_get_data.py" and test it  
Note: for the function to work, we must first have run 2a and 2e above,
in order to connect to server and define long, lat  
```{r}
# Source function (load the function into memory)
source_python("32_Function_get_data.py")

# Now we can use the function as it was an R function
get_data(selected_location$STATION_CODE[1],
         selected_location$LONGITUDE[1], 
         selected_location$LATITUDE[1],
         30)

```

### b. Pick several locations
```{r}

selected_locations <- data_stations %>%
  filter(STATION_CODE %in% c("HT4", "B07", "304"))

selected_locations

```

### c. Use purrr::map to get data for all locations
* map performs get_data() for every item sent to it through %>%
* In this case, we use map_df, which tells R to try to make a single data frame of the data
* inside map:
    - the tilde (~) says that the following should be regarded as a function 
    - the dot (.) symbolizes the input to the function
* 1:nrow(selected_locations) = 1:3 in this case
* So map runs 3 times, every time with the dot equal to 1,2,3  
* I.e, the first time, it runs
```
get_data(
    selected_locations$STATION_CODE[1],
    selected_locations$LONGITUDE[1], 
    selected_locations$LATITUDE[1],
    30)
```
The second time: 
```
get_data(
    selected_locations$STATION_CODE[2],
    selected_locations$LONGITUDE[2], 
    selected_locations$LATITUDE[2],
    30)
```
Etc.
    
```{r}

data_locations <- 1:nrow(selected_locations) %>%
  map_df(~get_data(
    selected_locations$STATION_CODE[.],
    selected_locations$LONGITUDE[.], 
    selected_locations$LATITUDE[.],
    30)
    )

# Make proper time variable
data_locations$Date <- as.POSIXct(
  data_locations$Time_num, 
  origin = "1970-01-01", tz = "UTC")

```


### d. Plot result  
```{r}
ggplot(data_locations, aes(x = Date, y = Temp, color = STATION_CODE)) +
  geom_line()

```

### e. Plot all results  
Note: requires tidyr::gather   
```{r}
# Reorganize the data using tidyr::gather
df <- data_locations %>%
  tidyr::gather("Variable", "Value", Temp:NO3)

# Plot all
ggplot(df, aes(x = Date, y = Value, color = STATION_CODE)) +
  geom_line() +
  facet_wrap(vars(Variable), scales = "free_y")

```

