---
title: "Read depth profile data from thredds server, as R function"
author: "DHJ"
date: "23 9 2019"
output: 
  html_document:
    keep_md: true
    toc_float: true

---

# Read and plot depth profile data from thredds server  

What we will do here, is to  
1. List variables (in order to find the variable names we need for step 2)   
2. Get all temperature and salinity data from given coordinates (all times, all depths)  
3. Plot those data (as profile plots and T-S plots)  
  
To read variable profiles for a single station, just a single line of code is needed, using the
function read_profile() - see section 2:  
```
my_data <- read_profile(10.5268, 59.0267, c('temp','salt'))
```
  
To read variable profiles for several stations, we need a data frame containing 3 columns (station names, longitudes,
and latitudes) - then this is used as input to the function read_profiles() - see section 4:  
```
my_data <- read_profiles(df_stations, "Station", c('N1_p','N3_n'))
```


## 0. Load libraries and functions  
```{r, results='hide', message=FALSE}
library(reticulate)
use_python("C:/WinPython/WPy64-3720/python-3.7.2.amd64/python.exe")   # my Python installation
library(ggplot2)
library(lubridate)
library(dplyr)
library(purrr)
library(mgcv)

# Functions Specifically for this 
source("34_Profile_data_R_functions.R")
source_python("34_Profile_data_Python_functions.py")

```


## 1. Get variable information   
* Use this function to find out which variables you can download and which names to use   
* you must use the name given in the first column (e.g. `salt`  for salinity). Sometimes it is hard to guess, such as `N1_p`for phosphate  
* This includes modelled biological variables (e.g. plankton density) and chemical ones (e.g. nitrate, phosphate). See part 4 for example    
* 'variable_info()' is in fact a Python function (loaded by 'source_python' above), but can be used by R as it was an ordinary R function!   
```{r}
variables <- variable_info()

# Will only show up in an interactive session:
View(variables)

```

## 2. Read data for a single position  
Reads data from the thredds server  
* In this example 
    - the location (10.52 E, 59.03 N) is the coordinates of Færder hard-bottom station   
    - the variables we ask for is temperature and salinity. Choose names from list above (section 2)
* When we don't specify server_url, it uses the default ('http://thredds.met.no/thredds/dodsC/metusers/arildb/MARTINI800_prov_v2.ncml')
* May take a minute or two    
```{r}
df1 <- read_profile(10.5268, 59.0267, c('temp','salt'))

# The next two lines does the same thing (read_profile just combines those two):
# May be useful for debugging in case of errors)
#   X <- read_profile_list(10.5268, 59.0267, c('temp','salt'))  # returns list
#   df1 <- profile_list2dataframe(X)                             # returns data.frame

```


## 3. Test plots  

### a. Profile plots  
May also smooth the surface before plotting - see script 33  
```{r}
ggplot(df1, aes(x = Time, y = Depth_mid, fill = temp, height = Depth_hi- Depth_lo)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue4", mid = "green", high = "red2", midpoint = 10)

ggplot(df1, aes(x = Time, y = Depth_mid, fill = salt, height = Depth_hi- Depth_lo)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue4", mid = "green", high = "red2", midpoint = 25)

```

### b. Tiles + contours, smoothed   
Here, using GAM with tensor product smooth (te) which is a bit slow   
```{r}
library(mgcv)

# Perform GAM and compute the predicted values from the GAM model
pred <- gam(temp ~ te(as.numeric(Time), Depth_mid, k = 20), data = df1) %>% predict()

# Add predicted values
df1$temp_gam = pred

# Plot predicted values, including contour lines
ggplot(df1, aes(x = Time, y = Depth_mid, fill = temp_gam, height = Depth_hi- Depth_lo)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue4", mid = "green", high = "red2", midpoint = 10) +
  geom_contour(aes(z = temp_gam), binwidth = 2, color = "white")

```

### c. T-S plot (temperatur vs. salinity) version 1 
```{r}
df1 <- df1 %>%
  mutate(Depth_binned = cut(Depth_mid, breaks = -c(0,10,20,40,90)))

ggplot(df1, aes(x = salt, y = temp)) +
  geom_point(size = rel(0.5)) +
  facet_grid(rows = vars(Depth_binned), cols = vars(month(Time)))

```

### d. T-S plot (temperatur vs. salinity) version 2  
```{r}
df1 <- df1 %>%
  mutate(Month_binned = cut(month(Time), breaks = c(0,3,6,9,12), 
                            labels = c("Jan-Mar", "Apr-Jun", "Jul-Sep", "Oct-Dec"))
         )

# Just to check:
# df1 %>% count(month(Time), Month_binned)

ggplot(df1, aes(x = salt, y = temp, col = Depth_mid)) +
  geom_point(size = rel(0.5)) +
  facet_wrap(vars(Month_binned))

```


## 4. Read data for several positions   
Input data is a data frame with names and positions  

### a. Get data  
* In this case, we want phosphate (N1_p) and nitrate (N3_n) from two stations defined by `df_stations`     
```{r}
df_stations <- data.frame(
  Station = c("B07 Tromøy", "HT4 Færder fyr"), 
  LONGITUDE = c(8.9443, 10.5268), 
  LATITUDE = c(58.5132, 59.0267)
  )

df2 <- read_profiles(df_stations, "Station", c('N1_p','N3_n'))

```

### b. Test plots    
* Note that the water is much shallower at Tromøy compared to Færder  
```{r}
ggplot(df2, aes(x = Time, y = Depth_mid, fill = N1_p, height = Depth_hi- Depth_lo)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue4", mid = "green", high = "red2", midpoint = 0.4) +
  facet_grid(cols = vars(ID)) +
  labs(title = "Phosphate (N1_p)")

ggplot(df2, aes(x = Time, y = Depth_mid, fill = N3_n, height = Depth_hi- Depth_lo)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue4", mid = "green", high = "red2", midpoint = 5) +
  facet_grid(cols = vars(ID)) +
  labs(title = "Nitrate (N3_n)")

```

